---
title: "Homework 4"
author: "Kristen Lee"
date: "2023-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Item 1
### Calculate Effect Size (Cohen's d and r)
#### Compare these results to the one from homework 2
Conduct a t-test to see whether life satisfaction depends upon the type of animal to which a person was married.
```{r}
goat <- read.csv ("https://blue.butler.edu/~rpadgett/ps310/Data/Goat_or_Dog.csv", 
      header= TRUE)
library(psych)
describe(goat)
cor.test(goat$life_satisfaction, goat$wife)
```

```{r}
library (effectsize)
goat$wife <- factor(goat$wife, levels=c(1,2), labels=c("goat","dog"))
cohens_d(goat$life_satisfaction, goat$wife, data=goat, pooled_sd=TRUE)
```

#### Two Sample t-test
```{r}
mod1<- t.test(life_satisfaction~wife, data = goat) #Welch
mod1
mod1$stderr #standard error difference
mod3<- t.test(life_satisfaction~wife, data = goat, var.equal=TRUE)
mod3
mod3$stderr
t2r(mod3$statistic,mod3$parameter)
```
```{r}
mod2<-lm(goat$life_satisfaction~goat$wife)## DV first
summary(mod2)
anova(mod2)

```

r = 0.63 and Cohen’s d = -1.57. The life satisfaction of men married to a dog wife was significantly higher than that of men married to a goat wife. This is similar to the results found in homework 2. 

## Item 2
### Do a paired t-test to see whether the hygiene scores on day 1 differ from those on day 3

```{r}
festive <- read.csv ("https://blue.butler.edu/~rpadgett/ps310/Data/DownloadFestival.csv", 
      header= TRUE)
festive <- na.omit(festive) 
```

#### Check the Descriptive Stats
```{r}
describe(festive$day1)
describe(festive$day3)
```

#### Run a paired t-test
```{r}
hygiene<-t.test(festive$day1, festive$day3, paired =TRUE, data = festive)
hygiene
hygiene$stderr
mod4<-cor.test(festive$day1,festive$day3) #paired samples correlation
mod4
```
#### Measure Effect Size
```{r message=FALSE, warning=FALSE}
library (effectsize)

cohens_d(hygiene)

t2r(hygiene$statistic,hygiene$parameter)
```

The hygiene scores significantly decreased from day 1 to day 3 of the music festival. 
