---
title: "Homework 3"
author: "Kristen Lee"
date: "2023-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1
```{r}
tea <- read.csv ("https://blue.butler.edu/~rpadgett/ps310/Data/tea_makes_you_brainy_716.csv", 
      header= TRUE)
library(psych)
describe(tea)
```

```{r}
library (olsrr)
mod5 <- lm(Cognitive_Function~Cups_of_Tea, data=tea)
ols_regress(mod5)
summary(mod5)
```

Based on the output we can see that the p-value = 0.38 so p < 0.05. So, we can reject the Ho which would state that there is no significant relationship between cognitive functioning and cups of tea that a person drinks. This model is able to significantly predict cognitive functioning based on cups of tea that a person drinks. 

To predict cognitive functioning if a person were to drink 10 cups of tea we can create the formula
cognitive functioning = 49.218 + 0.416(amount of tea in cups). So, the predicted cognitive functioning = 53.82 for 10 cups of tea.

## Task 4
```{r}
super <- read.csv ("https://blue.butler.edu/~rpadgett/ps310/Data/Supermodel.csv", 
      header= TRUE)
library(psych)
describe(super)
```

```{r}
library (olsrr)
mod6 <- lm(salary~age+years+beauty, data=super)
ols_regress(mod6)
summary(mod6)
```

Based on this output, age is a significant predictor of a model's salary. Looking at the beta, this is a positive relationship meaning that as a model's age increases, their salary increases. Years of experience or years of being a model is also a significant predictor, but it is a negative relationship. This means that as years goes up, the model's salary decreases. From this model we also see that beauty is not a significant predictor of a model's salary. 

```{r}
library (olsrr)
mod7 <- lm(salary~years, data=super)
ols_regress(mod7)

plot(super$age,super$salary)
plot(super$years,super$salary)
plot(super$beauty,super$salary)
```

The first finding was that years of experience increases, salary decreases, is not what one would expect especially since age has a positive relationship with salary. Here in this output when we do years of experience as the only predictor, the relationship is positive. Using multiple predictors can influence the output of a regression model. 


