---
title: "Homework 5"
author: "Kristen Lee"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Just Do the basic ANOVAs and Contrast effects, Effect sizes and interpret them.
```{r}
Teach <- read.csv("Teach.csv", header = TRUE)
library(psych)
Teach$group<-factor(Teach$group, ordered=TRUE, levels=c("Punish", "Indifferent", "Reward"))

describeBy(Teach$exam, Teach$group)
```

## Chapter 12 Task 1
### Fit a model with planned contrasts to test the hypotheses that: (1) reward results in better exam results than either punishment or indifference; and (2) indifference will lead to significantly better exam results than punishment.
```{r message=FALSE, warning=FALSE}
mod1 <- aov(exam ~ group, data=Teach)
summary (mod1)
```
From this ANOVA summary table output we can see that the significance value is less than 0.05 and we can say that teaching style has a significant effect on exam marks. We cannot conclude which group had what effect since we don't know which of them differed.  

##### Plotting the Means (This is not part of the homework)
```{r message=FALSE, warning=FALSE}
library(gplots)
plotmeans(exam~group, data=Teach)

mod2<-lm(exam~group, data=Teach)
summary(mod2)
```
##### Post HOC (This is not part of the homework)
```{r}
pairwise.t.test(Teach$exam, Teach$group, p.adj = "none")
TukeyHSD(mod1, conf.level = .95)
library(DescTools)
ScheffeTest(mod1, conf.level=.95)
```
We get differnt answers depending on which PostHOC test we use 
Scheffe is the most stringent 

### Contrasts
```{r message=FALSE, warning=FALSE}
library (emmeans)
emmeans(mod1, pairwise ~ group)

contrast1 <-  c(1,1,-2) # reward different from punish and indifferent 
contrast2 <-  c(1,-1,0) #indifferent and punish different from Controls
contrasts(Teach$group) = cbind(contrast1, contrast2) #cbind "oh i want you to bind these three things together"
contrasts(Teach$group) # this prints the contrast thingy with sig codes 
mod2 <- aov(exam ~ group, data=Teach) #Type 1 SS - not relevant for one-way ANOVA
summary.lm (mod2)
```
From this contrast output we can see that reward is significantly different than indifference and punishment (*p*-value < 0.05). We can also look at the descriptive statistics and see that mean exam score was significantly higher after reward when compared to indifferent or punish teaching styles.

This output also shows us that the indifferent teaching style produces significantly higher exam scores than the punishment teaching style (*p*-value < 0.05). From the descriptive statistics we see that indifference produces significantly higher exam scores than the punishment teaching style.

These results support the original hypotheses; (1) reward results in better exam results than either punishment or indifference; and (2) indifference will lead to significantly better exam results than punishment. We could conclude that teadchers or professors should reward their students, not punish them.

## Task 2
### Compute the effect sizes for the previous task.
```{r message=FALSE, warning=FALSE}
library(effectsize)
effectsize(mod1)
```
The Eta^2^ = 0.61 indicates that teaching style has a significant effect on exam scores.  
